% Metódy inžinierskej práce

\documentclass[10pt,twoside,slovak,a4paper]{article}

\usepackage[slovak]{babel}
\usepackage[IL2]{fontenc} 
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{url} 
\usepackage{hyperref} 

\usepackage{cite}
%\usepackage{times}

\pagestyle{headings}

\title{Potreba hier v mentálnom zdravie študenta\thanks{Semestrálny projekt v predmete Metódy inžinierskej práce, ak. rok 2022/23, vedenie: Vladimír Mlynarovič}}

\author{Andrii Ilkiv\\[2pt]
	{\small Slovenská technická univerzita v Bratislave}\\
	{\small Fakulta informatiky a informačných technológií}\\
	{\small \texttt{xilkiv@stuba.sk}}
	}

\date{\small 06. november 2022} 
\begin{document}
\maketitle
\begin{abstract}
\ldots
\end{abstract}
\section{Introduction}
Every mother in childhood said that video games are harmful and you should not play them. But is it the truth?  In my topic we are going to find out it. Most everyone now has access to smartphones, computers, PlayStations and that's why games are getting more and more popular day by day. People have been playing since childhood, games are natural. When a small child collects cubes, it develops the skill of construction, when children play social games, they try on roles, when a person plays chess, he learns to calculate actions. \newline However, the more difficult the game is, the more diligently a person plays - the greater the developing effect. A complex castle made of cubes - and the child understands that it is necessary to build carefully and thinking through the details. It's an interesting role that attracts the attention of others - and you come to understand that you need to take an active position in life. There are many moving combinations in chess.. In second part of my topic I will discuss it. In the third part of my topic I will tell you about effects of games.

\section{Discussion} \label{}
Nearly 80\texttt{\%} of gamers say that video games provide them with mental stimulation, relaxation and stress relief, other gamers such as athletes who live and breathe sports can often struggle with stress, depression and anxiety. Many gamers have  started playing video games to escape from something in real life. Perhaps they were bullied at school,  and when they came home, they were bullied on social networks. In such a situation, the only way out for a gamer is a video game. It is the only place where they can immerse themselves and  escape  the emotional pain of real life. The stereotype of the shy person who uses video games as an escape  is  not  at all what the average gamer looks like.\newline Those who played more video games were more likely to have good social skills in real life, perform better academically, and develop better relationships with other students in university because of the social and cooperative component of some types of games. Once upon a time I played a game. In it, I noticed that it is very easy for me to conduct various trade agreements and make deals. Before the game I did not have this experience, but in the game I felt that I got it. And of course I understood that was just a game, but the fact that I was the main trader of one of the top alliances was a reality. And all thoughts and efforts are also real from the point of view of the psyche. That is, the numbers are virtual, but the experience of management, calculations, analysis, negotiations, this experience is REAL, real efforts and the work of the psyche. Reasons to give a video game a try\cite{Geico.com}. \newline In other words, playing video games forces gamers to deeply stimulate multiple different areas of his or her brain, which leads to the development of a large handful of various beneficial tasks and dimensions of thinking. All computer games are games with rules, and all real objects in computer games are replaced by symbols that become more and more realistic with the development of technology. I would like to note one important difference between computer games and real ones: in a real game, the participants come up with the conditions of the game themselves, interact with each other, can change their behavior during the game, which allows for more flexible thinking, develops communication skills. Virtual games to a certain extent reduce these possibilities, since the game usually has a developer who puts a certain algorithm in it, which does not change radically during the game. 

\section{Effects} 
Gaming addiction is often seen as spending too much time playing video games,
and many people believe someone is addicted when they can't "turn off the controller".
There are two main types of video games and therefore two main types of video game addiction. Standard video games are usually single-player and have a clear goal or mission, such as rescuing a princess. Addiction in these games is often about completing that mission or achieving a high score. Another type of video game addiction is related to multiplayer online games. These games are played online with other people and are particularly addictive because they are usually endless. Gamers with this type of addiction enjoy creating and temporarily becoming an online character. They often build relationships with other online players as an escape from reality. For some, this community may be the place where they feel most accepted.

\newpage

\section{Depression and Anxiety: What Are Psychiatric Disorders?}
Psychiatry is the branch of medicine that focuses on mental, emotional, and behavioral disorders\cite{healthygamer.gg}. As such, the term, “psychiatric disorder,” refers to a broad range of problems that disturb a person's thoughts, feelings, behavior or mood.
Excessive video gaming is found to be associated with positive emotions and social relationships while playing. \newline However, problematic and excessive video gaming is also associated with negative emotions, low self-esteem, loneliness, and poor academic performance.Psychiatric disorders are incredibly complex and can affect individuals in numerous ways.

\subsection{What is Anxiety} \label{ina:este}
Anxiety is a common symptom gaming addicts experience. Anxiety can affect a person's life in many ways\cite{wired.com}. It comes in many forms, but often involves repetitive negative thoughts that are difficult to quiet. It can also be more than a failure to finish the game. Players can feel frustrated when they let their teammates down in multiplayer and feel lonely when playing solo.

\subsection{What is Depression } \label{ina:este}
The exact cause of depression is unknown. It may be caused by a combination of genetic, biological, environmental, and psychological factors. Everyone is different. You need to know that not all video games and gaming will lead to depression.\newline For example: suppose you play video games sporadically for an hour a day or so. Still, you have other activities in your life that fulfill you, and you have an overall stable life (a safe job, healthy relationships, and hobbies). In that case, video games will probably not cause depression. It's important to know that if you're struggling with a gaming problem and depression, you should focus on your real life. Implement strategies to reduce excessive gaming and address the underlying causes of your depression.



\bibliography{literatura}
\bibliographystyle{plain} 
\end{document}
